---
title: "Coronavirus Second Waves"
author: "Tyler DeGroff"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r, warning = FALSE, message = FALSE, echo = FALSE}
library(readxl)
library(dplyr)
library(zoo)
library(readr)
library(knitr)
```

# Import New York Times Coronavirus Data

```{r}
setwd("/Users/tylerdegroff/Documents/Github/NYTimes COVID-19 Data")

nytimes <- read.csv(
  file = "us-counties.csv",
  colClasses = c("Date", "factor", "factor", "character", "numeric", "numeric")
)

setwd("/Users/tylerdegroff/Documents/Github/Coronavirus")
```

**Source:** New York Times, The, Smith, M., Yourish, K., Almukhtar, S., Collins, K., Ivory, D., & Harmon, A. (2020, January–July). *Coronavirus (Covid-19) Data in the United States* [Cumulative counts of coronavirus cases in the United States, at the county level, over time (daily frequency).]. Github. https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv

\bigskip

```{r}
data <- nytimes[nytimes$state %in% c(
  "Alabama",              "Alaska",               "Arizona", 
  "Arkansas",             "California",           "Colorado", 
  "Connecticut",          "Delaware",             "District of Columbia", 
  "Florida",              "Georgia",              "Hawaii", 
  "Idaho",                "Illinois",             "Indiana", 
  "Iowa",                 "Kansas",               "Kentucky", 
  "Louisiana",            "Maine",                "Maryland",
  "Massachusetts",        "Michigan",             "Minnesota", 
  "Mississippi",          "Missouri",             "Montana",
  "Nebraska",             "Nevada",               "New Hampshire", 
  "New Jersey",           "New Mexico",           "New York",
  "North Carolina",       "North Dakota",         "Ohio", 
  "Oklahoma",             "Oregon",               "Pennsylvania",
  "Rhode Island",         "South Carolina",       "South Dakota", 
  "Tennessee",            "Texas",                "Utah", 
  "Vermont",              "Virginia",             "Washington", 
  "West Virginia",        "Wisconsin",            "Wyoming"
), ]

data <- droplevels(data)

data <- data %>% 
  within(., fips[county == "New York City"] <- 36999) # treats NYC as a county
```

# Import Census Bureau FIPS Code Metadata

```{r}
fips.state <- read_excel("fips.state.xlsx", skip = 4)
```

**Source:** United States Census Bureau. (2020, May 20). *2019 Census Bureau Region and Division Codes and State FIPS Codes* [Reference file for vintage 2019 Census Bureau state-level FIPS codes.]. United States Department of Commerce. https://www2.census.gov/programs-surveys/popest/geographies/2019/state-geocodes-v2019.xlsx

\bigskip

```{r}
fips.state <- fips.state %>%
  rename(fips.state = "State (FIPS)", state = "Name") %>%
  select(fips.state, state) %>%
  filter(fips.state != "00") %>%
  arrange(fips.state)
```

```{r}
fips.granular <- read_excel("fips.granular.xlsx", skip = 4)
```

**Source** United States Census Bureau. (2020, May 20). *2019 State, County, Minor Civil Division, and Incorporated Place FIPS Codes* [Reference file for vintage 2019 Census Bureau county-level FIPS codes.]. United States Department of Commerce. https://www2.census.gov/programs-surveys/popest/geographies/2019/all-geocodes-v2019.xlsx

\bigskip

```{r}
fips.county <- fips.granular %>%
  rename(
    fips.state = "State Code (FIPS)", 
    fips.county = "County Code (FIPS)",
    county = "Area Name (including legal/statistical area description)"
  ) %>%
  mutate(fips = as.numeric(paste0(
    as.character(fips.state), 
    as.character(fips.county)
  ))) %>%
  filter(fips.county != "000") %>%
  select(fips, fips.state, fips.county, county)
```

```{r}
fips <- merge(x = fips.county, y = fips.state, by = "fips.state", all.x = TRUE)
fips <- fips[complete.cases(fips[, "state"]), ]
```

# Import Census Bureau Population Estimates

```{r}
pop <- read_excel("pop.xlsx", skip = 3)
```

**Source:** United States Census Bureau. (2010–2019, April 1–July 1). *County Population Totals: 2010-2019* [Annual estimates of the county-level resident population, over time (annual frequency).]. United States Department of Commerce. https://www2.census.gov/programs-surveys/popest/tables/2010-2019/counties/totals/co-est2019-annres.xlsx

```{r, message = FALSE}
pop <- pop %>%
  rename(countyState = "...1", population = "2019") %>%
  filter(countyState != "United States") %>%
  select(countyState, population)
```

```{r}
fips <- fips %>% mutate(countyState = paste0(county, ", ", state))
pop <- merge(x = pop, y = fips, by = "countyState", all.x = TRUE)
```

# Aggregate and Treat NYC as Its Own County

```{r}
pop.nyc <- pop %>%
  filter(
    fips == 36005 | # Bronx County (Bronx)
    fips == 36047 | # Kings County (Brooklyn)
    fips == 36061 | # New York County (Manhattan)
    fips == 36081 | # Queens County (Queens)
    fips == 36085   # Richmond County (Staten Island)
  )

pop <- rbind(
  pop,
  data.frame(
    fips.state = "36",   # actual New York State state-level FIPS code
    fips.county = "999", # synthetic county-level FIPS code
    fips = "36999",      # synthetic FIPS code
    county = "New York City",
    state = "New York",
    countyState = "New York, New York",
    population = sum(pop.nyc$population)
  )
)
```

```{r}
data <- merge(
  x = data, 
  y = pop %>% select(fips, population), 
  by = "fips", 
  all.x = TRUE
)

data <- data %>%
  mutate(
    cases.percap = cases / population,
    deaths.percap = deaths / population
  )

data <- data[order(data$date, data$state, data$county), ]

data <- data %>%
  
  # mutate across dates by unique county/state combinations
  
  mutate(countyState = paste0(county, ", ", state)) %>%
  group_by(countyState) %>%
  
  mutate(
    cases.new = c(cases[1], diff(cases)),
    deaths.new = c(deaths[1], diff(deaths)),

    cases.new.percap = c(cases.percap[1], diff(cases.percap)),
    deaths.new.percap = c(deaths.percap[1], diff(deaths.percap)),
    
    cases.new.perM = cases.new.percap * 1000000,
    deaths.new.perM = deaths.new.percap * 1000000
  ) %>%
  
  mutate(
    cases.new.7dsma = rollmean(cases.new, k = 7, fill = 0, align = "right"),
    deaths.new.7dsma = rollmean(deaths.new, k = 7, fill = 0, align = "right"),
    
    cases.new.percap.7dsma = 
      rollmean(cases.new.percap, k = 7, fill = 0, align = "right"),
    deaths.new.percap.7dsma = 
      rollmean(deaths.new.percap, k = 7, fill = 0, align = "right"),
    
    cases.new.perM.7dsma = 
      rollmean(cases.new.perM, k = 7, fill = 0, align = "right"),
    deaths.new.perM.7dsma = 
      rollmean(deaths.new.perM, k = 7, fill = 0, align = "right"),
  ) %>%
  
  mutate(
    cases.active = rollsum(cases.new, k = 9, fill = 0, align = "right"),
    cases.active.percap = rollsum(cases.new.percap, k = 9, fill = 0, align = "right"),
    cases.active.perM = rollsum(cases.new.perM, k = 9, fill = 0, align = "right")
  )
```

```{r}
write_csv(data, "data.csv")
```

# Binomial Probability Distriubtion: Cumulative Density Function

\begin{align}
1 - P(q > x | n, p) &= 1 - {n \choose x} p^x \left( 1 - p \right)^{\left(n - x\right)}
\end{align}

```{r}
pbinom(
  q = 0, 
  size = 100, 
  prob = weighted.mean(x = wedding$cases.active.perM, w = wedding$guests) / 1000000,
  lower.tail = FALSE
)
```
